Cancer Applications of Choropleth Maps, and the Potential of Cartograms and Alternative Map Displays

### 2019-12-19 Response/Response.Rmd

---
from: Stephanie Kobakian
fromqualifications: BA
fromemail: stephanie.kobakian@monash.edu
fromphone: +61 433 699 797
fromsignature: sigfile.png
toname: Professor Peter Baade and Dr Susanna Cramb
opening: Dear Professor Peter Baade and Dr Susanna Cramb,
closing: Warm regards
spacing: 1.4
output: MonashEBSTemplates::letter
---

### 2019-12-19 review_article.Rmd
figure creation script

```{r setup, include=FALSE}
fig_path <- glue::glue("figs/")
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE,
                      fig.path = fig_path)
library(ggplot2)
library(cowplot)
library(png)
library(grid)
```

@@

```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Figure 1: A selection of choropleth cancer maps from online atlases that are publicly available. Maps of various countries were chosen: United Kingdom, Australia, Spain, USA, Canada, and display several different colour palettes and legends. These atllases are described in Table 1.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
library(cowplot)

# Import the images
plot1 <- png::readPNG('cancer_map_images/choropleths/21.choropleth.png')
plot2 <- png::readPNG('cancer_map_images/choropleths/3.choropleth.png')
plot3 <- png::readPNG('cancer_map_images/choropleths/25.choropleth.png')
plot4 <- png::readPNG('cancer_map_images/choropleths/15.choropleth.png')
plot5 <- png::readPNG('cancer_map_images/choropleths/7.choropleth.png')
plot6 <- png::readPNG('cancer_map_images/choropleths/23.choropleth.png')
plot7 <- png::readPNG('cancer_map_images/choropleths/26.choropleth.png')

ggdraw() +
  draw_plot(rasterGrob(plot1), 0, .75, 0.33, .25) +
  draw_plot(rasterGrob(plot2), 0.33, 0.75, 0.33, .25) +
  draw_plot(rasterGrob(plot3), 0.66, 0.5, .33, .5) +
  draw_plot(rasterGrob(plot4), 0, 0.5, .33, .25) +
  draw_plot(rasterGrob(plot5), 0.33, 0.5, .33, .25) +
  draw_plot(rasterGrob(plot6), 0, 0, .66, .5) +
  draw_plot(rasterGrob(plot7), .66, 0, .33, .5) +
  draw_plot_label(c("a", "b", "c", "d", "e", "f", "g"), 
    c(0.015, 0.31, 0.66, 0.015, 0.31, 0.014, 0.66), 
    c(0.99, 0.99, 0.99, 0.74, 0.74, 0.46, 0.46), size = 20)

ggsave(filename = "figures/choropleth_grid.png", 
  device = "png", width = 6, height = 5, dpi = 300)
```

@@

```{r ggcartograms, fig.cap = "Figure 2: Common alternatives to maps, showing the same information for the United States of America: (a) contiguous cartogram, (b) non-contiguous, shape-preserved cartogram, (c)  Dorling cartogram (non-contiguous), (d) hexagon tilemap (non-contiguous). Maps (a) - (c) are created by resizing and reshaping the states of the USA to match the 2015 population of the state. This provides a better sense of the extent of disease relative to the population in the country and can help ease losing information about physically small but population-dense states. Map creators give each state equal size and thus equal emphasis in (d) the hexagon tile map.", message=FALSE, warning=FALSE, out.width='100%'}
g1 <- png::readPNG("figures/ggcont.png")
g2 <- png::readPNG("figures/ggncont.png")
g3 <- png::readPNG("figures/ggdorl.png")
g4 <- png::readPNG("figures/gghexmap.png")
#knitr::include_graphics("figures/usa_grid.png")

load("figures/usa_legend.rda")

ggdraw() +
  draw_plot(rasterGrob(g1), 0, .55, 0.5,  0.45) +
  draw_plot(rasterGrob(g2), 0.5, 0.55, 0.5, 0.45) +
  draw_plot(rasterGrob(g3), 0.0, 0.1, 0.5, 0.45) +
  draw_plot(rasterGrob(g4), 0.5, 0.1, 0.5, 0.45) +
  draw_plot(usa_legend, 0, 0, 1, 0.1, scale = 2) +
  draw_plot_label(c("a", "b", "c", "d"), 
    c(0, .5, 0, 0.5), 
    c(1, 1, .5, .5), size = 20)
```

@@

```{r auscartograms, fig.cap = "Figure 4: Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tile map. The contiguous cartogram has expanded the highly populated areas while preserving the full shapes of rural areas. If it accurately sized areas by population, the country would be unrecognizable. The shape-preserved is unreadable due to the small area sizes. The Dorling cartogram presents all areas but many are difficult to compare. The hexagon tile map provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, out.width='100%', cache=FALSE, fig.width=8, fig.height=6, dpi=300}

aus1 <- png::readPNG("figures/aus_ggcont.png")
aus2 <- png::readPNG("figures/aus_ggncont.png")
aus3 <- png::readPNG("figures/aus_ggdorl.png")
aus4 <- png::readPNG("figures/aus_gghexmap.png")
load("aus_legend.rda")
load("figures/aus_legend.rda")

library(cowplot)
ggdraw(xlim = c(0,1), ylim = c(0,1.2)) +
  draw_plot(rasterGrob(aus1), 0, .65, 0.5, 0.50) +
  draw_plot(rasterGrob(aus2), 0.5, 0.65, 0.5, 0.50) +
  draw_plot(rasterGrob(aus3), 0.0, 0.1, 0.5, 0.55) +
  draw_plot(rasterGrob(aus4), 0.5, 0.1, 0.5, 0.55) +
  draw_plot(aus_legend, 0, 0, 1, .1) +
  draw_plot_label(c("a", "b", "c", "d"), 
    c(0, .5, 0, 0.5), 
    c(1.15, 1.15, .6, .6), size = 20)
```

@@

```{r interacting, fig.cap = "Figure 5: Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for [Globocan 2018: Cancer Today ](http://globocan.iarc.fr/Pages/Map.aspx), (ii) Menus for variable selection and zooming on [Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), (iii) Menus for choosing variables and countries in [The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M),  (iv) Tabs for different indicators and cancer types in [Global Cancer Map](http://globalcancermap.com/), (v) Menus and toggles for variable and subset selection in [United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html).", fig.width = 12, fig.height = 9}
i1 <- png::readPNG('cancer_map_images/interactivity/3.interactivity.png') # b in choro grid
i2 <- png::readPNG('cancer_map_images/interactivity/15.interactivity.png') # d in choro grid
i3 <- png::readPNG('cancer_map_images/interactivity/4.interactivity.png')
i4 <- png::readPNG('cancer_map_images/interactivity/5.interactivity.png')
i5 <- png::readPNG('cancer_map_images/interactivity/7.interactivity.png') # e in choro grid

ggdraw() +
  draw_plot(rasterGrob(i1), 0,  0.13,   .28,  .83) + #a
  draw_plot(rasterGrob(i2), 0.32, 0.56, .68, .4) + #b
  draw_plot(rasterGrob(i3), 0.32, 0.41, .68, .13) + #c
  draw_plot(rasterGrob(i4), 0.32, 0.14, .68, .25) + #d
  draw_plot(rasterGrob(i5), 0, 0, 1, .13) + #e
  draw_plot_label(c("i", "ii", "iii", "iv", "v"), 
    c(0.01, 0.3, 0.3, 0.3, 0.01), 
    c(0.99, 0.99, 0.57, 0.40, 0.15), size = 20)

ggsave(filename = "figures/supp_grid.png", device = "png", 
  width = 12, height = 9, dpi = 300)
```

```{r animating, fig.cap = "Two examples of advanced interactivity (and animation) in publicly available choropleth cancer maps: a. Linked maps and time-series line plots, with temporal animation in [Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC), b. A highly responsive magnifying glass on a map of [Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality).", fig.width = 12, fig.height = 9}
a1 <- png::readPNG("cancer_map_images/animation/23.animation.png")
a2 <- png::readPNG("cancer_map_images/animation/2.animation.png")

ggdraw() +
  draw_plot(rasterGrob(a1), 0, 0, 0.3, 1) +
  draw_plot(rasterGrob(a2), 0.3, 0, 0.7, 1) +
  draw_plot_label(c("I", "II"), 
    c(0, .3), c(1, 1), size = 20)
```

>>>>

```{r setup, include=FALSE}
fig_path <- glue::glue("figs/")
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE)
library(ggplot2)
library(cowplot)
library(png)
library(grid)
```

@@

```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Figure 1: A selection of choropleth cancer maps from online atlases that are publicly available. Maps of various countries were chosen: United Kingdom, Australia, Spain, USA, Canada, and display several different colour palettes and legends. These atllases are described in Table 1.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
knitr::include_graphics("figures/choropleth_grid.png")
```

@@

```{r ggcartograms, fig.cap = "Figure 2: Common alternatives to maps, showing the same information for the United States of America: (a) contiguous cartogram, (b) non-contiguous, shape-preserved cartogram, (c)  Dorling cartogram (non-contiguous), (d) hexagon tilemap (non-contiguous). Maps (a) - (c) are created by resizing and reshaping the states of the USA to match the 2015 population of the state. This provides a better sense of the extent of disease relative to the population in the country and can help ease losing information about physically small but population-dense states. Map creators give each state equal size and thus equal emphasis in (d) the hexagon tile map.", message=FALSE, warning=FALSE, out.width='100%'}
knitr::include_graphics("figures/usa_grid.png")
```

@@

```{r auscartograms, fig.cap = "Figure 4: Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tile map. The contiguous cartogram has expanded the highly populated areas while preserving the full shapes of rural areas. If it accurately sized areas by population, the country would be unrecognizable. The shape-preserved is unreadable due to the small area sizes. The Dorling cartogram presents all areas but many are difficult to compare. The hexagon tile map provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, out.width='100%', cache=FALSE, fig.width=8, fig.height=6, dpi=300}
knitr::include_graphics("figures/auscartograms.png")
```

@@

```{r interacting, fig.cap = "Figure 5: Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for [Globocan 2018: Cancer Today ](http://globocan.iarc.fr/Pages/Map.aspx), (ii) Menus for variable selection and zooming on [Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), (iii) Menus for choosing variables and countries in [The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M),  (iv) Tabs for different indicators and cancer types in [Global Cancer Map](http://globalcancermap.com/), (v) Menus and toggles for variable and subset selection in [United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html).", fig.width = 12, fig.height = 9}
knitr::include_graphics("figures/interacting.png")
```

```{r animating, fig.cap = "Two examples of advanced interactivity (and animation) in publicly available choropleth cancer maps: a. Linked maps and time-series line plots, with temporal animation in [Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC), b. A highly responsive magnifying glass on a map of [Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality).", fig.width = 12, fig.height = 7}
knitr::include_graphics("figures/animating.png")
```
